A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Beam Dynamics

### 2018-04-11 paper/theor-experiment.tex
adding results

\begin{align}
\text{min}  \quad & \text{rms}_{x, s_1}, \quad \text{rms}_{x, s_2} \label{eq:awa:p1}\\
& \text{rms}_{s, s_1}, \quad \text{rms}_{s, s_2} \label{eq:awa:p2}\\
& \varepsilon_{x,s_1}, \quad \varepsilon_{x,s_2} \label{eq:awa:p3} \\
& dE \label{eq:awa:p4} \\
\text{subject to} \quad & q = 40 \left[\text{nC}\right] \label{eq:awa:firstconstr}\\
\quad & \text{Volt}_{\text{Gun}} = 64\left[\text{MV/m}\right] \label{eq:awa:lastconstr}\\
\quad & \text{Volt}_{\text{Linac}} = 24-25\left[\text{MV/m}\right] \\
\quad & \sigma_x = \sigma_y = 9 \left[\text{mm}\right] \label{eq:awa:firstdvar}\\
\quad & \phi_{\text{linac}} =0^\circ \label{eq:awa:lastdvar}
\end{align}

@@

\subsubsection{AWA Optimization Results}
Given the parameters and variable boundaries in Section \ref{awaproblem}, 
four optimization experiments were done with varying GA parameters, see Table \ref{extable}.
From here on, we will reference each experiment as ex-1, ex-2, ex-3 and ex-4. 
They were each allowed to run for twenty four hours. 
Similar to the time step scan, the goal of this exercise was to determine which set of optimization
parameters strongly influence the results, and whether there was a time to solution
difference.

\begin{table}%[h!]
	\begin{center}
		\caption{Input Parameters for initial twenty four hour AWA optimization experiments. 
			The gene mutation probability was equal to the mutation probability (not shown) in all four experiments.}
		\label{extable}
		\begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} C C D }
			\hline\noalign{\smallskip}
			& Gene Mutation Probability & Recombination Probibility & Number of completed generations \\
			\noalign{\smallskip}\hline\noalign{\smallskip}
			ex-1 &  0.1  & 0.9  &  96 \\
			ex-2 &  0.3  & 0.7  &  81 \\
			ex-3 &  0.8  & 0.2  &  53 \\
			ex-4 &  0.01 & 0.09 &  95 \\
			\noalign{\smallskip}\hline
		\end{tabular*}
	\end{center}
\end{table}

After collection of all data, the Pareto front of each data set was plotted.

\vspace{0.2cm}
{\footnotesize\begin{verbatim}
OPTIMIZE, INPUT="tmpl/optLinac_40nC.tmpl",
	OUTPUT="optLinac_40nC",
	OUTDIR="results",
	OBJECTIVES = {rmss1, rmsx1, emitx1, de1, rmss2, rmsx2, emitx2},
	DVARS = {dv0, dv1, dv2, dv3, dv4, dv5, dv6, dv7},
	CONSTRAINTS = {c1, c2},
	INITIALPOPULATION=656,
	MAXGENERATIONS=200,
	NUM_MASTERS=1,
	NUM_COWORKERS=8,
	SIMTMPDIR="tmp",
	TEMPLATEDIR="tmpl",
	FIELDMAPDIR="fieldmaps",
	NUM_IND_GEN=328,
	GENE_MUTATION_PROBABILITY=0.8,
	MUTATION_PROBABILITY=0.8,
	RECOMBINATION_PROBABILITY=0.2;
QUIT;
\end{verbatim}}

>>>>

\begin{align}
\text{min}  \quad & \text{rms}_{x}\left(s = s_1\right), \quad \text{rms}_{x}\left(s = s_2\right) \label{eq:awa:p1}\\
& \text{rms}_{s}\left(s = s_1\right), \quad \text{rms}_{s}\left(s = s_2\right), \label{eq:awa:p2}\\
& \varepsilon_{x}\left(s = s_1\right), \quad \varepsilon_{x}\left(s = s_2\right), \label{eq:awa:p3} \\
& dE \label{eq:awa:p4} \\
\text{subject to} \quad & q = 40 \left[\text{nC}\right] \label{eq:awa:firstconstr}\\
\quad & \text{Volt}_{\text{Gun}} = 64\left[\text{MV/m}\right] \label{eq:awa:lastconstr}\\
\quad & \text{Volt}_{\text{Linac}} = 24-25\left[\text{MV/m}\right] \\
\quad & \sigma_x = \sigma_y = 9 \left[\text{mm}\right] \label{eq:awa:firstdvar}\\
\quad & \phi_{\text{linac}} =0^\circ \label{eq:awa:lastdvar}
\end{align}

@@

\subsubsection{AWA Optimization Results}
While the optimization problem and goals are well defined (Section \ref{awaproblem}), 
it is not clear what the best input parameters for the NSGA-II are.
These parameters include gene mutation probability, mutation probability, 
recombination probability, number of individuals (initial and and for later generations), and number of generations to complete. 
Given the beam line and design variables given in Section \ref{awaproblem}, 
four optimization experiments were done with varying NSGA-II parameters. 
Similar to the time step scan, 
the goal of this exercise was to determine which set of optimization
parameters strongly influence the results, 
and whether there was a time to solution difference.
From here on, we will reference each experiment as ex-1, ex-2, ex-3, and ex-4
as shown in Table \ref{extable}. 
\begin{table}%[h!]
	\begin{center}
		\caption{Input Parameters for initial twenty four hour AWA optimization experiments. 
			The gene mutation probability was equal to the mutation probability (not shown) in all four experiments. The number of individuals}
		\label{extable}
		\begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} C C D }
			\hline\noalign{\smallskip}
			& Gene Mutation Probability & Recombination Probibility & Number of completed generations \\
			\noalign{\smallskip}\hline\noalign{\smallskip}
			ex-1 &  0.1  & 0.9  &  96 \\
			ex-2 &  0.3  & 0.7  &  81 \\
			ex-3 &  0.8  & 0.2  &  53 \\
			ex-4 &  0.01 & 0.09 &  95 \\
			\noalign{\smallskip}\hline
		\end{tabular*}
	\end{center}
\end{table}

The initial number of individuals was fixed at 656, 
and the minimum number individuals in later generations was fixed at 328. 
These numbers we in part based on the architecture of the KNL's. 
Since each simulation takes 8 cores, and there are 64 cores per KNL node, 
we wanted a large population size that would fit evenly on these resources. 
Each experiment was allowed to run for twenty four hours, with 
a max generation limit of 100. 
We also reduced the six objectives to the four most important for quicker evaluation of these tests \nrnote{Q for Andreas, is it true less obj is faster}.
The objectives include: 
$\varepsilon_{x}\left(s = s_1\right)\text{, } \varepsilon_{x}\left(s = s_2\right)\text{, } \text{rms}_{s}\left(s = s_1\right)\text{, and }  \text{rms}_{s}\left(s = s_2\right)$. 
The OPAL input file for ex-3 is given as an example:

\vspace{0.2cm}
{\footnotesize \begin{verbatim}
	TITLE, STRING="ANL Optimisation";

	// Gun variables
	dv0: DVAR, VARIABLE="IBF",    LOWERBOUND=200.0, UPPERBOUND=500.0;
	dv1: DVAR, VARIABLE="IM",     LOWERBOUND=170.0, UPPERBOUND=260.0;
	dv2: DVAR, VARIABLE="GPHASE", LOWERBOUND=-30.0, UPPERBOUND=0.0;
	dv3: DVAR, VARIABLE="FWHM",   LOWERBOUND=1.5,   UPPERBOUND=10.0;

	// Objectives
	rmss1:  OBJECTIVE,EXPR="fabs(statVariableAt('rms_s',16.5))";
	emitx1: OBJECTIVE,EXPR="fabs(statVariableAt('emit_x',16.5))";	
	rmss2:  OBJECTIVE,EXPR="fabs(statVariableAt('rms_s',18.5))";
	emitx2: OBJECTIVE,EXPR="fabs(statVariableAt('emit_x',18.5))";

	OPTIMIZE, INPUT="tmpl/optLinac_40nC.tmpl",
	OUTPUT="optLinac_40nC",
	OUTDIR="results",
	OBJECTIVES = {rmss1, rmsx1, emitx1, de1, rmss2, rmsx2, emitx2},
	DVARS = {dv0, dv1, dv2, dv3, dv4, dv5, dv6, dv7},
	CONSTRAINTS = {c1, c2},
	INITIALPOPULATION=656,
	MAXGENERATIONS=200,
	NUM_MASTERS=1,
	NUM_COWORKERS=8,
	SIMTMPDIR="tmp",
	TEMPLATEDIR="tmpl",
	FIELDMAPDIR="fieldmaps",
	NUM_IND_GEN=328,
	GENE_MUTATION_PROBABILITY=0.8,
	MUTATION_PROBABILITY=0.8,
	RECOMBINATION_PROBABILITY=0.2;
	QUIT;
	\end{verbatim}}
\vspace{0.2cm}

After collection of the data for all four experiments, several metrics
were compared, including number of generations completed in twenty four hours,
location of Pareto fronts at $s_1$ and $s_2$, and the design variables at the Pareto front. From Table \ref{extable}, we clearly see ex-3 is significantly 
slower, as it evaluated about 30 to 40 fewer generations than the other experiments.
Perhaps this trade off would be acceptable if the Pareto front was significantly 
improved, but from Fig. \ref{expareto1}, we can see this is not the case.
The Pareto fronts at $s_1$, are nearly identical. We could expect
they would continue to converge given more time. 
When looking at the Pareto front for $s_2$, we see ex-4 has a slight 
edge over the others.  
\begin{figure}
	\centering
	\begin{subfigure}{0.49\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Report/ex-pareto1}
		\caption{Pareto fronts at $s_1$.}
		\label{exparetos1}
	\end{subfigure}
	\begin{subfigure}{0.49\textwidth}
	\centering
	\includegraphics[width=\textwidth]{Report/ex-pareto2}
	\caption{Pareto fronts at $s_2$.}
	\label{exparetos2}
	\end{subfigure}
	\caption{Comparison of Pareto fronts for initial experiments.}
	\label{expareto}
\end{figure}
